Macros for straightforward parallelism in Clojure

### 2016-03-23 paper/paper.tex
updated intro

\documentclass{sig-alternate}
\usepackage{minted}
\usemintedstyle{emacs}

@@

@@

% question: Can macros be used to create interesting and useful parallelism
% tools for recursive functions in clojure.
\section{Introduction}
% most important section in paper:
% - introduce the problem being solved
% - prove that this problem is interesting
% - why is this hard?
% - how does it differ from other approaches?
% - what are key results/limitations
Clojure is a lisp--like language running on the Java Virtual Machine.
The language's default immutable data structures and it's sophisticated STM system make it well suited for parallel and concurrent programming.
Additionally, because Clojure runs on the Java Virtual Machine, Clojure developers can take advantage of existing cross--platform parallelism libraries, such as Java's excellent \code{ExecutorService} framework, to write parallel code.

Taking advantage of Clojure's parallel potential is not entirely straightforward, however.
Interfacing with Java libraries directly is simple, but the interface code often feels slightly out of place when surrounded by other Clojure code.
The STM system is easy to use, but the STM constructs are often too low level to be immediately useful.
% for example, writing a parallel map with atoms would be a big pain

% is control flow the right word here? I feel like something else might be more appropriate.
% Recursion is control flow, parlet is control flow, so I guess it makes sense.
As a result, there are a variety of libraries designed to allow developers to take advantage of the parallelism and concurrency potential in the language.
Many of these library functions and builtins are data-parallel; they are designed to apply some sort of operation to a set of data.
This is often the kind of parallelism that developers desire, or are accustomed to identifying, but Clojure's nature as a functional language with immutable structures also makes it possible to exploit parallelism in the control flow of the code.

% even if it is expressed in control flow, it is still data parallelism probably
Using Clojure's macro system, we have implemented a few macros which allow developers to take advantage of Clojure's parallelism potential when their parallelism is more directly expressed in the control flow of their application.
We have shown that it is possible to attain noticeable degrees of parallelism with minimal code changes (with respect to serial code) through these macros, without the need for sophisticated dependency analysis (find OpenMP paper) often required in other parallelism systems.

@@

This paper is divided into $n$ sections. The first discusses the macros we've implemented, along with the appropriate use cases for each. In the second section, we present benchmarks demonstrating the cases in which the parallelism macros can (and cannot) improve performance of existing code.

@@

In this example, the expressions \mintcode{(foo value1)} and \mintcode{(foo value2)} are both evaluated as \code{ForkJoinTask}s in a \code{ForkJoinPool}.

@@

>>>>

\documentclass{sig-alternate}
% \usepackage[none]{hyphenat}
\usepackage{hyperref}

\usepackage{minted}
\usemintedstyle{emacs}

\bibliographystyle{plain}

@@

\author{
    Zmick, David\\
    \texttt{zmick2@illinois.edu}
}

\title{Macros for straightforward parallelism in Clojure}

@@

\maketitle

% question: Can macros be used to create interesting and useful parallelism
% tools for recursive functions in clojure.
\section{Introduction}
Clojure is a lisp--like language running on the Java Virtual Machine.
The language's default immutable data structures and it's sophisticated Software Transactional Memory (STM) system make it well suited for parallel programming.
Because Clojure runs on the Java Virtual Machine, Clojure developers can take advantage of existing cross--platform parallelism libraries, such as Java's excellent \code{ExecutorService} framework, to write parallel code.

% for example, writing a parallel map with atoms would be a big pain
Taking advantage of Clojure's parallel potential is not entirely straightforward, however.
STM has proven to be very successful as a clear construct for concurrent programming\cite{Jones2007a}, but these constructs are often too low level to be of much use to developers whose central concern is not parallelism\cite{Boehm2009}.

As a result, there are a variety of libraries designed to allow developers to take advantage of the parallelism potential in Clojure.\footnotemark{}\footnotemark{}\footnotemark{}\footnotemark{}\footnotemark{}
Many of these library functions and builtins are data-parallel; they are designed to apply some sort of operation to a set of data.
This is often the kind of parallelism that developers desire\cite{Okur2012a} but Clojure's nature as a functional language with immutable structures also makes it possible to easily exploit control parallelism (also known as task parallelism\cite{Andradea}).

\footnotetext[1]{\url{https://github.com/clojure/core.async}}
\footnotetext[2]{\url{https://github.com/TheClimateCorporation/claypoole}}
\footnotetext[3]{\url{https://clojuredocs.org/clojure.core/pmap}}
\footnotetext[4]{\url{http://clojure.org/reference/reducers}}
\footnotetext[5]{\url{https://github.com/aphyr/tesser}}

% even if it is expressed in control flow, it is still data parallelism probably
Using Clojure's macro system, we have implemented a set of macros which allow developers to take advantage of Clojure's parallelism potential when their existing code is written such that parallelism is exposed through control flow.
We have shown that it is possible to attain noticeable degrees of parallelism with minimal code changes (with respect to serial code), without the need for sophisticated dependency analysis  often required in other parallelism systems.

@@

@@

In this example, the expressions \mintcode{(foo value1)} and \mintcode{(foo value2)} are both evaluated as \code{ForkJoinTask}s in a \code{ForkJoinPool}\cite{Lea2000}.

@@

\bibliography{/home/dpzmick/Documents/Bibtex/senior-thesis.bib}
